---
title: "SRS_Assignment3"
output: pdf_document
date: "2023-03-23"
---
```{r}



data_m<-c(1, 1, 1, 15, 0, 3, 19, 54, 
                 4, 19, 62, 464, 76, 703, 1006)

#**** Coefficients: Indicator variables for the 4 Lists ******
# LA: Local Authority
# NGO: Non Governmental Organization
# PF: Police Force
# GO: Governmental Organization

LA<- c(rep(1,8),rep(0,7))
NGO<- c(rep(1,4),rep(0,4),rep(1,4),rep(0,3))
PF<-c(rep(c(1,0,1,0),3),c(1,0,1))
GO<-c(rep(c(1,1,0,0),3),c(1,1,0))

df<-data.frame(cbind(data_m,LA,NGO,PF,GO))
#***** 1.  Initial model ***********

model<-glm(data_m~LA+NGO+PF+GO, family=poisson(link="log"))

summary(model)
```


```{r}
#**** 2. Model with interaction terms: using all interaction terms *****

model <- glm(data_m~LA+NGO+PF+GO+(LA+NGO+PF+GO)^2, family=poisson(link="log"))

summary(model)

```




```{r}
#**** 3. Model with interaction terms: Adding terms based on AIC ****

step_model<-step(model)
summary(step_model)
```
The residual deviance is 11.341 for 7 degrees of freedom. The change in deviance is 6047.262-11.341 = 6035.921. 
Comparing this with the Ch-sq critical value for 14-7=7 degrees of freedom, we can reject the null hypothesis and conclude that our chosen model fits the data much better compared to an intercept only model

```{r}
anova(step_model)
```

# Final model
```{r}
final_model<-glm(data_m~LA+NGO+PF+GO+(LA*NGO)+(LA*PF)+(NGO*GO), family=poisson(link="log"))
summary(final_model)
```
```{r}
exp(summary(final_model)$coefficients[1,1])+sum(data_m)
```

```{r}
summary(final_model)$coefficients
```


```{r}
library(ggplot2)


# Create data frame for ggplot
df <- data.frame(fitted = step_model$fitted.values,
                 pearson_residual = pearson_residual)

# Pearson residuals plot
p1 <- ggplot(df, aes(x = fitted, y = pearson_residual)) +
  geom_point(color="steelblue")+
  geom_hline(aes(yintercept=0),color="#F8766D")+
  labs(x = "Fitted values (log scale)", y = "Pearson residuals") +
  theme_bw()

# QQ plot of Pearson residuals
p2 <- ggplot(df, aes(sample = pearson_residual)) +
  geom_qq(color="steelblue") +
  stat_qq_line(color="#F8766D") +
  labs(x = "Theoretical quantiles", y = "Pearson residuals") +
  theme_bw()

# Combine the plots
gridExtra::grid.arrange(p1, p2, ncol = 2)


```
```{r}
# Checcking fitted values against original
cbind(step_model$fitted.values,data_m,pearson_residual)
```



```{r}
prof<-profile(final_model,level=0.95)
```


# Profile Likelihood method for the confidence interval
```{r}

conf<-confint(profile(final_model,level=0.95))
conf
```

```{r}
exp(confint(profile(final_model,level=0.95))[1,1:2])+sum(data_m)
```


For the confidence intervals of the total population, exp(intervals for beta_0)+sum(data)
```{r}
exp(confint(profile(model_paper,level=0.95))[1,1:2])+sum(data)
```







